---
title: "Creating graphs in R with `ggplot2`"
author: "N Tabassum"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---
# notes for github
when saved, only then the files appear on git
# plot practice
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(here)
```

# Code for reading and plotting graphs
```{r}
ci_np <- read_excel(here("data","ci_np.xlsx"))
## First, tell R you’re using ggplot()
##Then, tell it the object name where variables exist (data = df_name)
##Next, tell it the aesthetics aes() to specify which variables you want to plot
##Then add a layer for the type of geom (graph type) with geom_*() - for example, geom_point() is a scatterplot, geom_line() is a line graph, geom_col() is a column graph, etc.
#let the first graph be an object
gg1 <- ggplot(data = ci_np, aes(x = year, y = visitors))

# now plotting different variations of gg1
##gg1 +
  ##geom_line()

# changing gg1 aesthetics
gg1 +
  geom_line(
    color = "brown",
    linetype = "dotted", ##changing line type
  )
```
Changing markers
```{r marker_plots}
gg1 + 
  geom_point(
    color = "cyan", #colors-go to Rcolors(https://r-graph-gallery.com/ggplot2-color.html)
             pch = 17, #characters-type?pch for guide
             size = 4,
             alpha = 1 #color channel
    )

```

## Mapping variables onto aesthetics
When we want to customize a graph element based on a variable’s characteristic or value, add the argument within aes() in the appropriate geom_*() layer

Example: Create a ggplot scatterplot graph where the size and color of the points change based on the number of visitors, and make all points the same level of opacity (alpha = 0.5). Notice the aes() around the size = and color = arguments.

```{r gg1_aes}
gg1 +
  geom_point(
    aes(color = visitors,
        size = visitors),
    alpha = 1
  )

```
Create a column plot of Channel Islands National Park visitation over time, where the fill color (argument: fill =) changes based on the number of visitors.

```{r geom_col_mapping to fill}
gg1 + 
  geom_col(aes(fill = visitors))
```
ggplot2 complete themes: While every element of a ggplot graph is manually customizable, there are also built-in themes (theme_*()) that you can add to your ggplot code to make some major headway before making smaller tweaks manually

```{r changing themes}
gg1 +
  geom_point(
    aes(size = visitors,
        color = visitors),
    alpha = 0.5
  ) +
  theme_minimal()
#removing grid lines, background and plot backgrounds
gg1 +
  geom_point(
    aes(size = visitors,
        color = visitors),
    alpha = 0.5
  ) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))
```

##Updating axis labels and titles
```{r axes}
gg1 +
  geom_line(color = 'violet') +
  #theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  labs(
    x = 'Year',
    y = "Annual park visitors"
  )

```

# Multi-series ggplot graphs 

```{r multiseries}
ca_np <- read_csv(here("data","ca_np.csv"))
gg2 <- ggplot(data = ca_np, aes(x = year, y = visitors, group = park_name))
gg2 + 
  geom_line()


```

# Faceting ggplot graphs (subplots)

```{r subplots}
gg2 +
  geom_line(show.legend = FALSE) +
    theme_light() + 
  labs(x = "year", y = "annual visitors") +
  facet_wrap(~ park_name)

```

# Exporting a ggplot graph with ggsave()
ggsave() will export the most recently run ggplot graph by default (plot = last_plot()), unless you give it the name of a different saved ggplot object.

```{r exporting}

ggsave(here("figures", "np_graph.jpg"), dpi = 180, width = 8, height = 7)
# Notice that a .jpg image of that name and size is now stored in the figures\ folder within your working directory. You can change the type of exported image, too (e.g. pdf, tiff, eps, png, mmp, svg).
```






