---
title: "Creating graphs in R with `ggplot2`"
author: "N Tabassum"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(here)
```

# Code for reading and plotting graphs
```{r}
ci_np <- read_excel(here("data","ci_np.xlsx"))
## First, tell R you’re using ggplot()
##Then, tell it the object name where variables exist (data = df_name)
##Next, tell it the aesthetics aes() to specify which variables you want to plot
##Then add a layer for the type of geom (graph type) with geom_*() - for example, geom_point() is a scatterplot, geom_line() is a line graph, geom_col() is a column graph, etc.
#let the first graph be an object
gg1 <- ggplot(data = ci_np, aes(x = year, y = visitors))

# now plotting different variations of gg1
##gg1 +
  ##geom_line()

# changing gg1 aesthetics
gg1 +
  geom_line(
    color = "brown",
    linetype = "dotted", ##changing line type
  )
```
Changing markers
```{r marker_plots}
gg1 + 
  geom_point(
    color = "cyan", #colors-go to Rcolors(https://r-graph-gallery.com/ggplot2-color.html)
             pch = 17, #characters-type?pch for guide
             size = 4,
             alpha = 1 #color channel
    )

```

## Mapping variables onto aesthetics
When we want to customize a graph element based on a variable’s characteristic or value, add the argument within aes() in the appropriate geom_*() layer

Example: Create a ggplot scatterplot graph where the size and color of the points change based on the number of visitors, and make all points the same level of opacity (alpha = 0.5). Notice the aes() around the size = and color = arguments.

```{r gg1_aes}
gg1 +
  geom_point(
    aes(color = visitors,
        size = visitors),
    alpha = 1
  )

```
Create a column plot of Channel Islands National Park visitation over time, where the fill color (argument: fill =) changes based on the number of visitors.

```{r geom_col_mapping to fill}
gg1 + 
  geom_col(aes(fill = visitors))
```
ggplot2 complete themes: While every element of a ggplot graph is manually customizable, there are also built-in themes (theme_*()) that you can add to your ggplot code to make some major headway before making smaller tweaks manually

```{r changing themes}
gg1 +
  geom_point(
    aes(size = visitors,
        color = visitors),
    alpha = 0.5
  ) +
  theme_minimal()
#removing grid lines, background and plot backgrounds
gg1 +
  geom_point(
    aes(size = visitors,
        color = visitors),
    alpha = 0.5
  ) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))
```

##Updating axis labels and titles




